The \vspec{} Collection: A suite of utilities to model spectroscopic phase curves of 3D exoplanet atmospheres in the presence of stellar variability.

### 2024-08-20 src/tex/ms.tex
Clean up article
Move figs around, add inkscape acknowledgement
add orcid IDS
Add orcid IDS
Merge response to Veselin's comments

% Define document class
\documentclass[linenumbers,preprint,authoryear]{elsarticle}
\usepackage{showyourwork}
\usepackage[version=4]{mhchem}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{natbib}

@@

\author[1,2,3,4,5]{Ted M Johnson\corref{cor1}}
\ead{ted.johnson@unlv.edu}

\author[1,2,3]{Cameron Kelahan}

\author[3]{Avi M. Mandell}

\author[3]{Ashraf Dhahbi}

\author[3]{Tobi Hammond}

\author[3]{Thomas Barclay}

\author[3]{Veselin B. Kostov}
\author[3]{Geronimo L. Villanueva}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/ld_transit.pdf}
    \caption{
        {\bf Right}: Transit light curves at 1 ${\rm \mu m}$ for a Lambertian (i.e. no limb darkening) star and one with significant limb darkening \citep[$u_1=0.5,\,u_2=0.15$,][]{espinoza2015}. In each case a 1 $R_e$ planet transits a 0.12 $R_\odot$ ($T_{\rm eff} = 3000$ K) star with an observation cadence of 1 minute. Scattered points show the observation with simulated noise, considering a single wavelength channel from JWST/NIRSpec G395H binned to a resolving power of 270 and a target distance of 5 pc.
        {\bf Left}: Stellar surface maps including translucent masks to show limb darkening and transit. The Lambertian star (top) does not exhibit any limb darkening, and its disk has constant surface brightness. However, the limb-darkened star (bottom) is much brighter at disk-center than it is near the limb. The dark circle on the leftmost limb of each star is the region occulted by the planet when it is $0.7^\circ$ from mid-transit.
    }
    \script{ld_transit.py}
    \label{fig:ld_transit}
\end{figure*}

In the case of a transiting geometry, \texttt{vspec-vsm} is responsible for computing the properties of the portion of the stellar disk that is occulted. It does this by first projecting both the transiting planet and the visible part of the surface orthographically as they would be viewed by a distant observer. It then selects all the points that are near enough to the transit to be occulted. The code then iterates through each of those points and computes the fraction of it covered by the occulting circle using a 2D numerical integral. The weight of each point is determined by the geometric projected area as well as a quadratic limb darkening law \citep[see][]{espinoza2015}. This treatment allows the user to specify two parameters $u_1$ and $u_2$ that describe the flux variation as a function of angle from disk-center. A caveat of this method is that there is no dependence on wavelength. A better treatment would vary the temperature of the stellar model as a function of $\mu$; we simply apply a gray scaling factor to the flux. Figure \ref{fig:ld_transit} compares transit light curves for two different sets of limb darkening parameters.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/jax_v_scipy.pdf}
    \caption{
        Speed comparison between the JAX and SciPy implementations of the GridPolator interpolation backend. Both interpolators use the builtin \vspec{} grid, with \teff ranging from 2800 to 3300K, and wavelengths from 5-12 ${\rm \mu m}$ with $R=100$. SciPy is most efficient for a small number of evaluations, with a constant cost per call. For a large number of evaluations, however, the speed gained via JAX's JIT compilation offsets the cost of performing that compilation.
    }
    \script{jax_v_scipy.py}
    \label{fig:jax_v_scipy}
\end{figure}

@@

To compute the surface temperature we first compute the temperature at the equator as a function of longitude. This is done by integrating Equation (10) of \citet{cowan2011}
\begin{equation}
    \frac{d \tilde{T}}{d \Phi} = \frac{1}{\epsilon}(\max{(\cos{\Phi}, 0)} - \tilde{T}^4)
\end{equation}
Where $\tilde{T}$ is the dimensionless temperature defined as $T/T_0$, or the temperature normalized by a local fiducial value. $\Phi$ is the angle from the substellar point, so we numerically integrate it from $-\pi$ to $\pi$ using one of several different schemes depending on the value of $\epsilon$. For $0 < \epsilon < 1$ we adopt a reflecting initial value problem (IVP) scheme, where we set $\tilde{T}=1$ at $\Phi=\pi$, integrate to $-\pi$, then use the last value as the initial value for an integration from $-\pi$ to $\pi$. Note that for extremely small values of $\epsilon$ (e.g. $10^{-6}$), the integration is slow, and will fail if $\epsilon=0$; we recommended a lower limit of $\epsilon=10^{-4}$. When $1\le \epsilon < 10$ we solve the integral using a boundary value problem (BVP) scheme. For each of these two methods we employ the SciPy \citep{virtanen2020} implementation: \texttt{solve\_ivp()}, employing a 5th order Runga-Kutta scheme \citep{dormand1980} and \texttt{solve\_bvp()}, based on an earlier MATLAB implementation \citep{kierzenka2001}, respectively. When $\epsilon \ge 10$ it is more efficient to use the analytic approximation given by Equation (A2) in \citet{cowan2011}. In Figure \ref{fig:cowan_curves} we show results for several different values of $\epsilon$.

@@

\subsection{Phase Curve}
\begin{figure}[t]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/kempton23.pdf}
    \caption{
        A simulation of a MIRI phase curve for GJ 1214b, binned by wavelength; compare to real data published by \citet{kempton2023}(Extended Data Figure 1). While the addition of stellar variability does have a visible effect on the light curve, it is much smaller than the thermal emission produced by the exoplanet.
        }
    \script{kempton23.py}
    \label{fig:gj1214b}
\end{figure}

@@

The figures in this article were created using the Matplotlib \citep{hunter2007}, NumPy \citep{harris2020}, and Cartopy \citep{metoffice2010} libraries.

>>>>

% Define document class
\documentclass[linenumbers,5p,twocolumn,authoryear]{elsarticle}
\usepackage{showyourwork}
\usepackage[version=4]{mhchem}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{xcolor}
\usepackage{orcidlink}

@@

\author[1,2,3,4,5]{Ted M Johnson\,\orcidlink{0000-0002-1570-2203}\corref{cor1}}
\ead{ted.johnson@unlv.edu}

\author[1,2,3]{Cameron Kelahan\,\orcidlink{0009-0002-0075-8041}}

\author[3]{Avi M. Mandell\,\orcidlink{0000-0002-8119-3355}}

\author[3]{Ashraf Dhahbi\,\orcidlink{0009-0009-0200-6821}}

\author[3]{Tobi Hammond\,\orcidlink{0009-0002-5756-9778}}

\author[3]{Thomas Barclay\,\orcidlink{0000-0001-7139-2724}}

\author[3]{Veselin B. Kostov\,\orcidlink{0000-0001-9786-1031}}
\author[3]{Geronimo L. Villanueva\,\orcidlink{0000-0002-2662-5776}}

@@

\begin{figure*}[h]
    \centering
    \includegraphics[width=\textwidth]{figures/ld_transit.pdf}
    \caption{
        {\bf Right}: Transit light curves at 1 ${\rm \mu m}$ for a Lambertian (i.e. no limb darkening) star and one with significant limb darkening \citep[$u_1=0.5,\,u_2=0.15$,][]{espinoza2015}. In each case a 1 $R_e$ planet transits a 0.12 $R_\odot$ ($T_{\rm eff} = 3000$ K) star with an observation cadence of 1 minute. Scattered points show the observation with simulated noise, considering a single wavelength channel from JWST/NIRSpec G395H binned to a resolving power of 270 and a target distance of 5 pc.
        {\bf Left}: Stellar surface maps including translucent masks to show limb darkening and transit. The Lambertian star (top) does not exhibit any limb darkening, and its disk has constant surface brightness. However, the limb-darkened star (bottom) is much brighter at disk-center than it is near the limb. The dark circle on the leftmost limb of each star is the region occulted by the planet when it is $0.7^\circ$ from mid-transit.
    }
    \script{ld_transit.py}
    \label{fig:ld_transit}
\end{figure*}

In the case of a transiting geometry, \texttt{vspec-vsm} is responsible for computing the properties of the portion of the stellar disk that is occulted. It does this by first projecting every visible point (i.e. those for which $\mu > 0$, $\mu$ is the cosine of the angle from disk-center) orthographically onto an $x$-$y$ plane where the stellar disk is represented by the unit circle. The planet, in this projection, is represented by a circle centered at $(a/R_*\,\sin{\phi},~ a/R_*\,\cos{\phi}\,\cos{i})$ with a radius of $R_p/R_*$, for semi-major axis $a$, stellar radius $R_*$, planetary radius $R_p$, phase past midtransit $\phi$, and inclination $i$. It then iterates through each point on the surface, computing the distance in projected coordinates to the center of the occulting disk. Also, for each point, an effective radius is calculated in the projected coordinates
\begin{equation}
    r_{\rm eff} = 2\mu\sqrt{\Omega/(4\pi\,\text{[sr]})}
\end{equation}
where $\Omega$ is the solid angle of each point. Any point for which the distance is less than $R_p/R_* + 2r_{\rm eff}$ from the occulting disk's center is flagged as possibly interacting with the planet. The majority of these points are completely covered by the planet, and those that are less than $R_p/R_* - 2r_{\rm eff}$ from the center of the occultation are immediately marked as such. The remaining points -- those near the edge of the occultation -- are subdivided into a $100\times 100$ point meshgrid. A 2D numerical integral is performed over this grid to compute the fraction of each point that is inside the occulting disk.

 When computing the disk-integrated flux (including flux removed via a transit) the weight of each point is determined by the geometric projected area as well as a quadratic limb darkening law \citep[see][]{espinoza2015}. We treat limb darkening as a weight factor on the contribution of each point to the disk-integrated spectrum that depends on $\mu$. Points that are part of spots and faculae are treated the same as the photospheric points -- their respective weights are increased if they lie near the center and decreased if they lie on the limb. This treatment allows the user to specify two parameters $u_1$ and $u_2$ that describe the flux variation as a function of angle from disk-center. A caveat of this method is that there is no dependence on wavelength. A better treatment would vary the temperature of the stellar model as a function of $\mu$; we simply apply a gray scaling factor to the flux. Figure \ref{fig:ld_transit} compares transit light curves for two different sets of limb darkening parameters.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/jax_v_scipy.pdf}
    \caption{
        Speed comparison between the JAX and SciPy implementations of the GridPolator interpolation backend. Both interpolators use the builtin \vspec{} grid, with \teff ranging from 2800 to 3300K, and wavelengths from 5-12 ${\rm \mu m}$ with $R=100$. SciPy is most efficient for a small number of evaluations, with a constant cost per call. For a large number of evaluations, however, the speed gained via JAX's JIT compilation offsets the cost of performing that compilation.
    }
    \script{jax_v_scipy.py}
    \label{fig:jax_v_scipy}
\end{figure}

@@

To compute the surface temperature we first compute the temperature at the equator as a function of longitude. This is done by integrating \citet{cowan2011} Equation (10).
\begin{equation}
    \frac{d \tilde{T}}{d \Phi} = \frac{1}{\epsilon}(\max{(\cos{\Phi}, 0)} - \tilde{T}^4)
\end{equation}
Where $\tilde{T}$ is the dimensionless temperature defined as $T/T_0$, or the temperature normalized by a local fiducial value. $\Phi$ is the angle from the substellar point, so we numerically integrate it from $-\pi$ to $\pi$ using one of several different schemes depending on the value of $\epsilon$. For $0 < \epsilon < 1$ we adopt a reflecting initial value problem (IVP) scheme, where we set $\tilde{T}=1$ at $\Phi=\pi$, integrate to $-\pi$, then use the last value as the initial value for an integration from $-\pi$ to $\pi$. Note that for extremely small values of $\epsilon$ (e.g. $10^{-6}$), the integration is slow, and will fail if $\epsilon=0$; we recommended a lower limit of $\epsilon=10^{-4}$. When $1\le \epsilon < 10$ we solve the integral using a boundary value problem (BVP) scheme. For each of these two methods we employ the SciPy \citep{virtanen2020} implementation: \texttt{solve\_ivp()}, employing a 5th order Runga-Kutta scheme \citep{dormand1980} and \texttt{solve\_bvp()}, based on an earlier MATLAB implementation \citep{kierzenka2001}, respectively. When $\epsilon \ge 10$ it is more efficient to use the analytic approximation given by Equation (A2) in \citet{cowan2011}. In Figure \ref{fig:cowan_curves} we show results for several different values of $\epsilon$.

@@

\subsection{Phase Curve}
\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/kempton23.pdf}
    \caption{
        A simulation of a MIRI phase curve for GJ 1214b, binned by wavelength; compare to real data published by \citet{kempton2023}(Extended Data Figure 1). While the addition of stellar variability does have a visible effect on the light curve, it is much smaller than the thermal emission produced by the exoplanet.
        }
    \script{kempton23.py}
    \label{fig:gj1214b}
\end{figure}

@@

The figures in this article were created using the Matplotlib \citep{hunter2007}, NumPy \citep{harris2020}, and Cartopy \citep{metoffice2010} libraries. The top panel of Figure \ref{fig:fac_struct} was created using the open-source vector graphics software Inkscape\footnote{\url{https://inkscape.org/}}.
